---
title: "BIOSTAT620_HW2"
author: "Bulun Te"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### 1(a)

B

### 1(b)

C

### 1(c)

AD

### 1(d)

D

### 1(e)

AB


## Problem 2


```{r,include=FALSE}
library(tidyverse)
library(patchwork)
library(knitr)
library(kableExtra)

# Setting default language as English
Sys.setenv(LANGUAGE = "en")
Sys.setlocale("LC_TIME", "en_US.UTF-8")

data <- read.csv("BulunTe_IphoneMonitor.csv", header = TRUE)

data = data %>% mutate(proportion = Social.ST.min/Total.ST.min,
                       duration_per_use = Total.ST.min/Pickups)

data$Date = as.Date(data$Date,format="%Y/%m/%d")
data$weekday = (weekdays(data$Date)  %in% c( "Monday", 
                                             "Tuesday", 
                                             "Wednesday", 
                                             "Thursday", 
                                             "Friday")) %>% as.numeric() 
data$semester = (data$Date > date("2024-01-09")) %>% as.numeric()
data$semester_weekday = data$semester * data$weekday

data = data %>% filter(Date<="2024-01-26")

```



```{r}

data = data %>% mutate(
  Total.ST.min_lag1 = lag(Total.ST.min,1),
  Social.ST.min_lag1 = lag(Social.ST.min,1)
) %>% filter(!is.na(Total.ST.min_lag1))

data %>% head()


```



### (2a)

```{r}

library(systemfit)

eq1 = Total.ST.min ~ 1 + Total.ST.min_lag1 + weekday + semester
eq2 = Social.ST.min ~ 1 + Social.ST.min_lag1 + weekday + semester
eqs = list(eq1, eq2)

fit_2a = systemfit(eqs,method="SUR",data=data)

fit_2a %>% summary()

```



### (2b)

From the estimation, it is shown that at the significance level of 0.05, only the coefficient of `semster` is significant in the model for Total Screen time.The coefficient of the 'semester' variable in equation 1 (eq1) is estimated to be 69.984660 with a p-value of 0.036785, which indicates that, all else being equal, the daily total screen time is expected to be about 70 minutes greater on days after January 10 compared to days before January 10.

The reasons that cause only the coefficient of `semester` to be significant in the model for Total Screen time might be as follows:

1. With only 26 observations and 22 degrees of freedom, the sample size may be too small to detect significant effects.

2. The model might be missing some key variables or non-linear terms that are important in explaining the variation in the dependent variables.

3. As the Social Screen time and total screen time all increase in the semster, there might be multicollinearity between the semester and lag 1 of the screen times.












