---
title: "BIOSTAT620_HW2"
author: "Bulun Te"
date: "2024-03-05"
output: 
  pdf_document:
    latex_engine: xelatex
  html_document:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### 1(a)

B

### 1(b)

C

### 1(c)

AD

### 1(d)

D

### 1(e)

AB


## Problem 2


```{r,include=FALSE}
library(tidyverse)
library(patchwork)
library(knitr)
library(kableExtra)

# Setting default language as English
Sys.setenv(LANGUAGE = "en")
Sys.setlocale("LC_TIME", "en_US.UTF-8")

data <- read.csv("BulunTe_IphoneMonitor.csv", header = TRUE)

data = data %>% mutate(proportion = Social.ST.min/Total.ST.min,
                       duration_per_use = Total.ST.min/Pickups)

data$Date = as.Date(data$Date,format="%Y/%m/%d")
data$weekday = (weekdays(data$Date)  %in% c( "Monday", 
                                             "Tuesday", 
                                             "Wednesday", 
                                             "Thursday", 
                                             "Friday")) %>% as.numeric() 
data$semester = (data$Date > date("2024-01-09")) %>% as.numeric()
data$semester_weekday = data$semester * data$weekday

data = data %>% filter(Date<="2024-01-26")

```



```{r}

data = data %>% mutate(
  Total.ST.min_lag1 = lag(Total.ST.min,1),
  Social.ST.min_lag1 = lag(Social.ST.min,1)
) %>% filter(!is.na(Total.ST.min_lag1))

data %>% head()


```



### (2a)

```{r}

library(systemfit)

eq1 = Total.ST.min ~ 1 + Total.ST.min_lag1 + weekday + semester
eq2 = Social.ST.min ~ 1 + Social.ST.min_lag1 + weekday + semester
eqs = list(eq1, eq2)

fit_2a = systemfit(eqs,method="SUR",data=data)

fit_2a %>% summary()

```



### (2b)

From the estimation, it is shown that at the significance level of 0.05, only the coefficient of `semster` is significant in the model for Total Screen time.The coefficient of the 'semester' variable in equation 1 (eq1) is estimated to be 69.984660 with a p-value of 0.036785, which indicates that, all else being equal, the daily total screen time is expected to be about 70 minutes greater on days after January 10 compared to days before January 10.

The reasons that cause only the coefficient of `semester` to be significant in the model for Total Screen time might be as follows:

1. With only 26 observations and 22 degrees of freedom, the sample size may be too small to detect significant effects.

2. The model might be missing some key variables or non-linear terms that are important in explaining the variation in the dependent variables.

3. As the Social Screen time and total screen time all increase in the semster, there might be multicollinearity between the semester and lag 1 of the screen times.


### (2c)

```{r}

library(car)
hypothesis <- c("eq1_semester = 0", "eq2_semester = 0")
wald_test <- linearHypothesis(fit_2a, hypothesis)
wald_test

```


The test yielded an F-statistic of 2.5414 with a corresponding p-value of 0.09026. At the significance level of 0.05, we fail to reject the null hypothesis that the coefficients of the `semester` variable in both equations are jointly equal to zero. This indicates that the data doesn`t have sufficient evidence to confirm semester has a significant effect on the screen times.



## Problem 3


### (a)


By randomly assigning participants to either take A drug or take B drug, it ensures that confounders are likely be ditributed equally among the groups. In this way, any confounders have an equal chances of affecting each treatment arms. This means that unobserved confounders in the $\epsilon_i$ term are likely to be independent of the treatment assignment $X_i$. 

Because of random assignment, the treatment variable $X_i$ is independent of other variables that are not observed in the study. In this way, the treatment effect of drugs are independent of the outcome variable.

### (b)

As $X_i$ is coded in {1,-1}, thus, treatment effect for drug a is $\beta_1$ and treatment effect for drug b is $-\beta_1$. 


### (c)

$\text{Cov}(Y, X) = \text{Cov}(\beta_0 + \beta_1 X + \varepsilon, X) = \beta_1 \text{Var}(X) +\text{Cov}(\varepsilon, X)= \beta_1 \text{Var}(X)$

Therefore, without adding confounders,

$\hat{\beta}_1 = \frac{\text{Cov}(Y, X)}{\text{Var}(X)}$

After adding confounder Z, as by the randomization, Z is independent with X. Thus, we have

$\text{Cov}(Y, X)= \text{Cov}(\beta_0 + \beta_1 X + \beta_2 Z + \varepsilon, X)= \beta_1 \text{Var}(X) + \beta_2 \text{Cov}(Z, X) + \text{Cov}(\varepsilon, X)=\beta_1 \text{Var}(X) + 0 + 0$

$\hat{\beta}_1 = \frac{\text{Cov}(Y, X)}{\text{Var}(X)}$

Therefore, by adding confounder, the estimate of $\beta_1$ is still unbiased.



### (d)

$\text{ATE} = \mathbb{E}\left[Y_A - Y_B\right] = \mathbb{E}\left[Y_A\right] - \mathbb{E}\left[Y_B\right]$

As X  is randomized,X is independent with outcome Y. Therefore,


$\mathbb{E}\left[Y_A\right] = \mathbb{E}\left[Y | X=1\right] = \beta_0 + \beta_1, \quad \mathbb{E}\left[Y_B\right] = \mathbb{E}\left[Y | X=-1\right] = \beta_0 - \beta_1$

Thus, $\text{ATE} = \mathbb{E}\left[Y_A - Y_B\right] = \mathbb{E}\left[Y_A\right] - \mathbb{E}\left[Y_B\right] = 2\beta_1$

